Learning and comparing functional connectomes across subjects

### 2013-01-01 paper.tex
iter

Support vector machines\footnote{See \cite{kloppel2008}
for a gentle introduction to support vector machines.} (classification and regression) are popular
algorithms for predictive modeling as they work on a large class of problems 
and do not require fine parameter setting.
However, in general learning algorithms do require the specification of
parameters, the values of which will impact the quality of the results.
These parameters may be specified based on domain specific knowledge,
determined using an analytical approach, or set using
cross validation to optimize the prediction accuracy. In latter case 
to avoid biasing measures of prediction accuracy, a
nested cross-validation scheme \cite{pereira2009} is required, assessing 
prediction on left-out data not used to set the parameters.
%
Selecting features most relevant to the model can help prediction
accuracy and model interpretation \cite{pereira2009}. One feature selection strategy is 
to filter out features
based on their statistical relationship with the variable of interest
using a univariate statistical criterion such as t-test, Pearson's
correlation, or mutual information.  A drawback of this approach is that
since the filters are typically insensitive to multivariate
information, they may exclude features that would be important to the
model. Note that feature selection must be performed inside the
cross-validation loop, \emph{i.e.}\ on the train data, excluding the test
data.

@@

The interpretation of the predictive model can be complicated and
requires insight into the mechanisms
underlying the modeling approach. For linear classifiers, 
classifier weights are similar to the weights of a ordinary regression.
If features are appropriately
standardized prior to training, the magnitude of weights can be
interpreted to reflect the relative importance of the feature to the
model, but interpreting how the feature differs between classes or
relates to a continuous variable is more complicated, particularly given
the multivariate nature of their involvement\footnote{Note that there is
typically no statistical framework for thresholding these weights, and
meaningless features are not always set to zero (this is particularly the
case for $\ell_2$ regularized algorithms such as SVM and ridge
regression)}. For classification methods
such as Fisher's linear discriminant analysis (LDA), which maximally
separates class means given their standard deviation, feature weights are
representative of the differences between class means. For support vector
classification, classifier weights are derived from observations that are
the most similar between classes, and hence represents the border between
classes. To draw conclusions on the features used for prediction, it is
perhaps most reasonable to adapt a conservative interpretation in which
predictive models is used to identify the most relevant features that are
later tested in followon experiments better suited to elucidating their
relationship to the variable of interest.

>>>>

Support vector machines\footnote{See \cite{kloppel2008}
for a gentle introduction to support vector machines.} (classification and regression) are popular
algorithms for predictive modeling as they work on a large class of problems 
and do not require fine parameter setting.
However, in general learning algorithms do require the specification of
parameters, the values of which will impact the quality of the results.
These parameters may be specified based on domain specific knowledge,
determined using an analytical approach, or set using
cross validation to optimize the prediction accuracy. In latter case 
to avoid biasing measures of prediction accuracy, a
nested cross-validation scheme \cite{pereira2009} is required, assessing 
prediction on left-out data not used to set the parameters.
%
Selecting features most relevant to the model can help prediction
accuracy and model interpretation \cite{pereira2009}; for instance with a
functional connectome by highlighting the connections that underpin the
prediction. One feature selection strategy is 
to filter out features
based on their statistical relationship with the variable of interest
using a univariate statistical criterion such as t-test, Pearson's
correlation, or mutual information.  A drawback of this approach is that
since the filters are typically insensitive to multivariate
information, they may exclude features that would be important to the
model. Note that feature selection must be performed inside the
cross-validation loop, \emph{i.e.}\ on the train data, excluding the test
data.

@@

The interpretation of the predictive model can be complicated and
requires insight into the mechanisms
underlying the modeling approach. For linear classifiers, 
classifier weights are similar to the weights of a ordinary regression.
If features are appropriately
standardized prior to training, the magnitude of weights can be
interpreted to reflect the relative importance of the feature to the
model, but interpreting how the feature differs between classes or
relates to a continuous variable is more complicated, particularly given
the multivariate nature of their involvement\footnote{Note that there is
typically no statistical framework for thresholding these weights, and
meaningless features are not always set to zero (this is particularly the
case for $\ell_2$ regularized algorithms such as SVM and ridge
regression)}. For classification methods
such as Fisher's linear discriminant analysis (LDA), which maximally
separates class means given their standard deviation, feature weights are
representative of the differences between class means. For support vector
classification, classifier weights are derived from observations that are
the most similar between classes, and hence represents the border between
classes. To draw conclusions on the features used for prediction, it is
perhaps most reasonable to adapt a conservative interpretation in which
predictive models is used to identify the most relevant features that are
later tested in follow-up experiments better suited to elucidating their
relationship to the variable of interest.
